---
title: "Centroid Detection Service"
author: "Brian Maitner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Centroid Detection Service}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## About the Centroid Detection Service

The Centroid Detection Service (CDS) is a tool for automated identification of potential geographic centroids in coordinate data.



# Using the CDS

The CDS takes in a dataframe of latitude and longitude points and spits out a bunch of stuff about whether they are likely to represent political division centroids.


```{r }
library(CDS)

#First, we'll grab an example dataset containing two columns: latitude and longitude.
coordinates <- read.csv("https://bien.nceas.ucsb.edu/bien/wp-content/uploads/2020/10/cds_testfile.csv", stringsAsFactors = FALSE)

#Take a quick look at the input format
head(coordinates)

#Run the query
CDS_output <- CDS(coordinates = coordinates)

#Take a look at what is returned
tail(CDS_output,10)


```

# Information returned by the CDS

There are a lot of columns returned by the CDS, some of which you may not care about, some of which you (presumably) do.  This includes:

 - Information on political divisions coordinates fall within.

 - Information on both absolute and relative distance to centroids at 3 levels of political divisions.

 - Opinions on whether a given coordinate is likely to be a centroid.
 
 - Information on whether a coordinate is not valid.

```{r}

#Information on political divisions coordinates fall within
tail(CDS_output[c("country","state","county")],n = 10)

#Information on absolute and relative distances (analagous columns for state and county levels)
tail(CDS_output[c("country_cent_dist","country_cent_dist_relative","country_cent_type","country_cent_dist_max","is_country_centroid")],10)

#Are the points likely centroids?
tail(CDS_output[c("is_country_centroid","is_state_centroid","is_county_centroid")],10)

#Are there any probems with the coordinates?
head(CDS_output["latlong_err"])

```
# CDS metadata
For information on the current version of the CDS, use the function CDS_version.

```{r}

CDS_version()

```
